DATA 621 - Business Analytics and Data Mining

### 2020-09-22 HW1/Drafts/HW1.Rmd

## 1. DATA EXPLORATION (Amber)
Describe the size and the variables in the moneyball training data set. Consider that too much detail will cause a manager to lose interest while too little detail will make the manager consider that you aren’t doing your job. Some suggestions are given below. Please do NOT treat this as a check list of things to do to complete the assignment. You should have your own thoughts on what to tell the boss. These are just ideas.

>>>>

## 1. DATA EXPLORATION (Amber)
Describe the size and the variables in the moneyball training data set. Consider
that too much detail will cause a manager to lose interest while too little
detail will make the manager consider that you aren't doing your job. Some
suggestions are given below. Please do NOT treat this as a check list of things
to do to complete the assignment. You should have your own thoughts on what to
tell the boss. These are just ideas.

### 2020-09-22 HW1/Drafts/HW1_draft_kellogg.Rmd
Cleanup folder by moving to draft

---
title: 'DATA 621---Business Analytics and Data Mining'
subtitle: 'Fall 2020---Group 2---Homework #1'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "9/27/2020"
output:
  pdf_document:
    toc: TRUE
    toc_depth: 2
  html_document: default
---

@@

# CODE APPENDIX
The code chunks below represent the R code called in order during the analysis.
They are reproduced in the appendix for review and comment.
```{r appendix, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```
```{r setup}
```
```{r loadData}
```
```{r wins}
```
```{r pctNull}
```
```{r negFactors}
```
```{r posFactors}
```
```{r summaryInfo}
```
```{r fewOutliers}
```
```{r manyOutliers}
```
```{r otherOutliers}
```
```{r corrPlot}
```
```{r outliers_1}
```
```{r iqr_removal_SO}
```
```{r iqr_removal_H}
```
```{r outlier_boxplot}
```
```{r impute_train}
```
```{r singles}
```
```{r make_slg_obp}
```
```{r slugging_compare}
```
```{r baseModel}
```
```{r cooks_distance}
```
```{r model2C+S}
```
```{r model2}
```
```{r model3}
```
```{r performance}
```
<!--```{r eval_set}
```-->
```{r predictions}
```

>>>>

<<<<<<< HEAD:HW1/HW1_draft_kellogg.Rmd
---
title: 'DATA 621---Business Analytics and Data Mining'
subtitle: 'Fall 2020---Group 2---Homework #1'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "9/27/2020"
output:
  pdf_document:
    toc: TRUE
    toc_depth: 2
  html_document: default
---

Studying the coefficients of the model suggest that winning is in favor if the
team batting hits more doubles, triples and home runs. Moreover, increase in the
number of stolen bases, and a decrease in caught steals, double plays, error,
and walks allowed would all lead to a win for the batting team. It is noteworthy
that the model suggests that a decrease in single hits by batter and an increase
in strikeouts by batters which seems counter intuitive. But these variables were
kept because when a batter steps to the plate, the player is more likely to
strike out than to get a hit. Trying to hit the ball out of the park will come
with strikeouts but it will also increase the chances of hitting home runs (even
1B, 2B, 3B), and that is pretty good exchange that most teams are willing carry
out.

# CODE APPENDIX
The code chunks below represent the R code called in order during the analysis.
They are reproduced in the appendix for review and comment.
```{r appendix, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```
```{r setup}
```
```{r loadData}
```
```{r wins}
```
```{r pctNull}
```
```{r negFactors}
```
```{r posFactors}
```
```{r summaryInfo}
```
```{r fewOutliers}
```
```{r manyOutliers}
```
```{r otherOutliers}
```
```{r corrPlot}
```
```{r outliers_1}
```
```{r iqr_removal_SO}
```
```{r iqr_removal_H}
```
```{r outlier_boxplot}
```
```{r impute_train}
```
```{r singles}
```
```{r make_slg_obp}
```
```{r slugging_compare}
```
```{r baseModel}
```
```{r cooks_distance}
```
```{r model2C+S}
```
```{r model2}
```
```{r model3}
```
```{r performance}
```
<!--```{r eval_set}
```-->
```{r predictions}
```
=======
---
title: 'DATA 621---Business Analytics and Data Mining'
subtitle: 'Fall 2020---Group 2---Homework #1'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "9/27/2020"
output:
  pdf_document:
    toc: TRUE
    toc_depth: 2
  html_document: default
---

@@

# CODE APPENDIX
The code chunks below represent the R code called in order during the analysis.
They are reproduced in the appendix for review and comment.
```{r appendix, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```
```{r setup}
```
```{r loadData}
```
```{r wins}
```
```{r pctNull}
```
```{r negFactors}
```
```{r posFactors}
```
```{r summaryInfo}
```
```{r fewOutliers}
```
```{r manyOutliers}
```
```{r otherOutliers}
```
```{r corrPlot}
```
```{r outliers_1}
```
```{r iqr_removal_SO}
```
```{r iqr_removal_H}
```
```{r outlier_boxplot}
```
```{r impute_train}
```
```{r singles}
```
```{r make_slg_obp}
```
```{r slugging_compare}
```
```{r baseModel}
```
```{r cooks_distance}
```
```{r model2C+S}
```
```{r model2}
```
```{r model3}
```
```{r performance}
```
<!--```{r eval_set}
```-->
```{r predictions}
```
>>>>>>> 84033f5f2f8e65ba0e7413efc0a539e933834f6f:HW1/Drafts/HW1_draft_kellogg.Rmd

### 2020-09-22 HW2/HW2-Group2.rmd

```{r}
library(tidyverse)

urlRemote  = "https://raw.githubusercontent.com/"
pathGithub = "aadler/DT621_Fall2020_Group2/master/HW2/"
fileTrain   = "classification-output-data.csv"
data = paste0(urlRemote, pathGithub, fileTrain) %>% read.csv()
```

@@

* class: the actual class for the observation  
* scored.class: the predicted class for the observation (based on a threshold of 0.5)  
* scored.probability: the predicted probability of success for the observation

```{r data_table}
class_data = data %>% select(class, pred_class = scored.class, pred_prob = scored.probability) 
class_data %>% head() %>% knitr::kable(caption = "Preview of the Dataset")
table(class_data[,1:2]) %>% knitr::kable(caption = "Counts of the Classification Data")
```

@@

* Accuracy: the proportion of the total number of predictions that were correct.  
* Positive Predictive Value or Precision: the proportion of positive cases that were correctly identified.  
* Negative Predictive Value: the proportion of negative cases that were correctly identified.  
* Sensitivity or Recall: the proportion of actual positive cases which are correctly identified.   
* Specificity: the proportion of actual negative cases which are correctly identified.

@@

@@

>>>>

```{r dataImport}
library(tidyverse)
library(pkr)
library(ggplot2)
library(caret)

# download data
urlRemote  = "https://raw.githubusercontent.com/"
pathGithub = "aadler/DT621_Fall2020_Group2/master/HW2/"
fileTrain   = "classification-output-data.csv"
data = paste0(urlRemote, pathGithub, fileTrain) %>% read.csv()
```

@@

* **class**: the actual class for the observation  
* **scored.class**: the predicted class for the observation (based on a threshold of 0.5)  
* **scored.probability**: the predicted probability of success for the observation

Use the table() function to get the raw confusion matrix for this scored dataset. Make sure you understand the output. In particular, do the rows represent the actual or predicted class? The columns?

```{r data_table}

class_data = data %>% select(class, pred_class = scored.class, pred_prob = scored.probability)

# raw confusion matrix
table("Predicted"= data$scored.class, "Actual"= data$class)

#class_data %>% head() %>% knitr::kable(caption = "Preview of the Dataset")
#table(class_data[,1:2]) %>% knitr::kable(caption = "Counts of the Classification Data")
```

**Rows represent the predicted value and columns represent the actual value.**

* 119 records were predicted to be 0 that were actually 0. (True negative)
* 30 records were predicted to be 0 that were actually 1. (False negative)
* 5 records were predicted to be 1 that were actually 0. (False positive)
* 27 records were predicted to be 1 that were actually 1. (True positive)

@@

* **Accuracy**: the proportion of the total number of predictions that were correct.  
* **Positive Predictive Value or Precision**: the proportion of positive cases that were correctly identified.  
* **Negative Predictive Value**: the proportion of negative cases that were correctly identified.  
* **Sensitivity or Recall**: the proportion of actual positive cases which are correctly identified.   
* Specificity: the proportion of actual negative cases which are correctly identified.

@@

**What are the bounds on the F1 score? Show that the F1 score will always be between 0 and 1. (Hint: If 0 < 𝑎 < 1 and 0 < 𝑏 < 1 then 𝑎𝑏 < 𝑎.)**

We know that both precision and sensitivity must be between 0 and 1.

* If these metrics are 0, it means that no true positives are identified. 
* If precision = 1, this means that there are no false positives. In other words, we didn't identify any records as being 1 that are actually 0. 
* If sensitivity = 1, this means that there are no false negatives. In other words, we didn't identify any records as 0 that are actually 1.

We know that the maximum possible value of the F1 score occurs when both precision and sensitivity are 1. If we plug these values into our F1 score equation, we can see that the upper bound is 1:
$F1\ score = \frac{2*1*1}{1+1} = \frac{2}{2} = 1$

Similarly, we know that the minimum possible value of the F1 score occurs when either precision or sensitivity equals 0 because the numerator in the equation ends up being 0. We can see that our lower bound is therefore 0:
$F1\ score = \frac{2*0*1}{1+0} = \frac{0}{1} = 0$

Therefore, the bounds for an F1 score are [0,1].

### ROC Curve
Function `roc_score` that generates an ROC curve from a data set with a true classification column (class in our example) and a probability column (scored.probability in our example). Your function should return a list that includes the plot of the ROC curve and a vector that contains the calculated area under the curve (AUC). Note that I recommend using a sequence of thresholds ranging from 0 to 1 at 0.01 intervals.

```{r roc, echo=TRUE}
# function for roc score 
roc_score <- function(df){
  if (is.data.frame(df) == FALSE) {
    return("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {

    seqOver <- seq(0,1,0.01)
    finalFrame <- data.frame(threshold=double(), 
                           tpr=double(),
                           fpr=double(),
                           stringsAsFactors=FALSE)

    for (i in seqOver){
      df2 <- df %>%
        mutate(tp = ifelse(class == 1 & pred_class == 1, 1, 0),
               tn = ifelse(class == 0 & pred_class == 0, 1, 0),
               fp = ifelse(class == 0 & pred_class == 1, 1, 0),
               fn = ifelse(class == 1 & pred_class == 0, 1, 0)) %>%
        filter(pred_prob <= i)

      if (nrow(df2) > 0){
        tpr <- sum(df2$tp)/ (sum(df2$tp)+sum(df2$fn))
        fpr <- sum(df2$fp)/ (sum(df2$fp)+sum(df2$tn))
        finalFrame <- rbind(finalFrame,data.frame(i, tpr, fpr))
      }
    }

    roc_plt <- ggplot(data=finalFrame, aes(x=fpr, y=tpr)) +
      geom_line() +
      ggtitle("ROC Curve") +
      xlab("False Positive Rate") +
      ylab("True Positive Rate") +
      geom_abline(intercept = 0, slope = 1)

    return(roc_plt)
  }

}

roc_final <- roc_score(class_data)
roc_final$plot_env$roc_plt

```

### Data Metrics

@@

### Caret Package

```{rpROC Package
